### 2021-04-27 papers/jin-thesis/Chapters/3Methods.tex
THESIS DONE (first fully complete draft)

The entire model and simulation code can be found at \url{https://github.com/kiatvonj/dynein_walk}, where the model and simulation were coded using both Python and C++. The both bound Monte Carlo was coded on Python due to the calculations being less computationally rigorous, while the time dependent one-bound Brownian dynamics was coded on C++ to utilize faster computation speed.

\section{Simulation}
%\textit{Picking Random angles to create random distribution of dynein configurations. Make sure to clarify that the simulation is split into two parts, the monte carlo both bound and brownian one bound. It is done this way in order for MC to generate ensemble of data rather quickly.}\\
%\textit{Include a flow chart.}
To generate an ensemble of steps, we use the Monte Carlo method to simulate a single step at a time, as opposed to a walk. This simplifies the code into two parts (the Monte Carlo both bound and the Brownian dynamics one bound) that cycle repeatedly after the model takes a step. The simulation starts in the both bound state, where the Monte Carlo randomly picks a both bound configuration and calculates the total energy. If the configuration successfully unbinds, then we use the positions of the domains as initial conditions for the one bound simulation and execute the Brownian dynamics. Once the one bound dynein diffuses onto the microtubule and rebinds, the simulation repeats the cycle and generates a completely different both bound configuration for another step. However, this method comes with a caveat: we must initialize the distance between the binding domains, $L$, to ensure that our ensemble covers the whole sample space of possible both bound configurations. This forces the measured probabilities to be a function of $L$ and will be addressed later in the Data Analysis section (Section \ref{sec:DataAna}). The algorithm for the both bound process is shown below:

@@

\begin{enumerate}
	\setcounter{enumi}{4}
	\item Calculate the total energy of the configuration by summing the energy of each domain (using Equation (\ref{eqn:energy})) and the relative unbinding probability from a Boltzmann distribution,
	\begin{equation}
		E_{total, i}=\sum_{k}\frac{1}{2}c_k(\theta_k-\theta_{k,eq})^2,
	\end{equation}
	\begin{equation}
		P_{ub} \propto \rho_{ub}e^{-\beta E_{total, i}}\kappa.
	\end{equation}
	Then, add a term to the partition function,
	\begin{equation}
		Z=\sum_{i}e^{\beta E_{total, i}}.
	\end{equation}
	Here, $i$ refers to the specific both bound configuration and $k$ is the domain. We also defined $\kappa$ to normalize the relative unbinding probability to keep it less than 1.

@@

\begin{enumerate}
	\setcounter{enumi}{6}
	\item If dynein unbinds, initialize the one bound state with the both bound configuration, run the Brownian dynamics simulation of the step, and wait until it rebinds.
\end{enumerate}

@@

\begin{enumerate}
	\setcounter{enumi}{7}
	\item Collect statistics about the final position and one bound time. Repeat algorithm from step 2. 
\end{enumerate}

@@

A more in-depth explanation and rigorous flow chart for the one bound Brownian dynamics simulation can be found here \cite{Capek2017, }.

%\section{Time Evolution}
%\textit{Simulate over a delta t during one bound and every iteration check for rebinding. How are the probabilities of binding and unbinding affected by the time.}

@@

\begin{figure}[hbt!]
	\centering
	\includegraphics[width=0.3\columnwidth]{../../plots/burgess-model-figure.pdf}
	\includegraphics[width=0.5\columnwidth]{../../plots/grotjahn-model-figure.pdf}%
	\caption[Experimental Pictures of Dynein for Model Parameters]{\textbf{Experimental Pictures of Dynein for Model Parameters} \textit{Left:} Micrograph of both bound dynein with model domains to identify post-stroke equilbrium angles. The arrow indicates 15 nm \cite{Burgess2003}.  \textit{Right:} Cryo-electron tomograph of one bound dynein for pre-stroke equilbrium angles \cite{grotjahn2018cryo}.} 
	\label{fig:ParamsPics}
\end{figure}

@@

Despite simulating many combinations of parameters, the number of free variables still produce many degrees of freedom. This makes it difficult to test the entire parameter space and truly invesitage the relationships between different sets of parameters. We approach this problem by individually attributing the parameters to a specific data set from exerpiment. For instance, the exponential unbinding constant, $C$, is fitted specifically to Figure (\ref{fig:trailingbias}). We use time data (\textit{cite paper for time data}\cite{}) to fit the unbinding rate constant, $k_{ub}$, because the inverse yields the average time it takes to unbind, i.e. the both bound time. The rebinding rates, $k_a$ and $k_b$, are harder to optimize because they directly correlate to step length. The ability for dynein to take long steps is dependent on how long dynein diffuses above the microtubule. Therefore, the rebinding rates have a much larger parameter space due to the flexibility of optimizing according to the one bound time or step length. Since we are interested in Yildiz's step lengths, we match these rates to the linear regression from Figure (\ref{fig:YildizCorrelation}) and Yildiz's step length plots. The last parameters are the spring constants. The spring constants also influence the step length because they dictate the strength of the restoring force. We fit the spring constants last to control the variance of step lengths.

>>>>

The entire model and simulation code can be found at \url{https://github.com/kiatvonj/dynein_walk}, where the model and simulation were coded using both Python and C++. The both-bound Monte Carlo was coded on Python due to the calculations being less computationally rigorous, while the time dependent one-bound Brownian dynamics was coded on C++ to utilize faster computation speed.

\section{Simulation}
%\textit{Picking Random angles to create random distribution of dynein configurations. Make sure to clarify that the simulation is split into two parts, the monte carlo both-bound and brownian one-bound. It is done this way in order for MC to generate ensemble of data rather quickly.}\\
%\textit{Include a flow chart.}
To generate an ensemble of steps, we use the Monte Carlo method to simulate a single step at a time, as opposed to a walk. This simplifies the code into two parts (the Monte Carlo both-bound and the Brownian dynamics one-bound) that cycle repeatedly after the model takes a step. The simulation starts in the both-bound state, where the Monte Carlo randomly picks a both-bound configuration and calculates the total energy. If the configuration successfully unbinds, then we use the positions of the domains as initial conditions for the one-bound simulation and execute the Brownian dynamics. Once the one-bound dynein diffuses onto the microtubule and rebinds, the simulation repeats the cycle and generates a completely different both-bound configuration for another step. However, this method comes with a caveat: we must initialize the distance between the binding domains, $L$, to ensure that our ensemble covers the whole sample space of possible both-bound configurations. This forces the measured probabilities to be a function of $L$ and will be addressed later in the Data Analysis section (Section \ref{sec:DataAna}). The algorithm for the both-bound process is shown below:

@@

\begin{enumerate}
	\setcounter{enumi}{4}
	\item Calculate the total energy of the configuration by summing the energy of each domain (using Equation (\ref{eqn:energy})) and the relative unbinding probability from a Boltzmann distribution,
	\begin{equation}
		E_{total, i}=\sum_{k}\frac{1}{2}c_k(\theta_k-\theta_{k,eq})^2,
	\end{equation}
	\begin{equation}
		P_{ub} \propto \rho_{ub}e^{-\beta E_{total, i}}\kappa.
	\end{equation}
	Then, add a term to the partition function,
	\begin{equation}
		Z=\sum_{i}e^{\beta E_{total, i}}.
	\end{equation}
	Here, $i$ refers to the specific both-bound configuration and $k$ is the domain. We also defined $\kappa$ to normalize the relative unbinding probability to keep it less than 1.

@@

\begin{enumerate}
	\setcounter{enumi}{6}
	\item If dynein unbinds, initialize the one-bound state with the both-bound configuration, run the Brownian dynamics simulation of the step, and wait until it rebinds.
\end{enumerate}

@@

\begin{enumerate}
	\setcounter{enumi}{7}
	\item Collect statistics about the final position and one-bound time. Repeat algorithm from step 2. 
\end{enumerate}

@@

A more in-depth explanation and rigorous flow chart for the one-bound Brownian dynamics simulation can be found here \cite{Capek2017, }.

%\section{Time Evolution}
%\textit{Simulate over a delta t during one-bound and every iteration check for rebinding. How are the probabilities of binding and unbinding affected by the time.}

@@

\begin{figure}[hbt!]
	\centering
	\includegraphics[width=0.3\columnwidth]{../../plots/burgess-model-figure.pdf}
	\includegraphics[width=0.5\columnwidth]{../../plots/grotjahn-model-figure.pdf}%
	\caption[Experimental Pictures of Dynein for Model Parameters]{\textbf{Experimental Pictures of Dynein for Model Parameters} \textit{Left:} Micrograph of both-bound dynein with model domains to identify post-stroke equilbrium angles. The arrow indicates 15 nm \cite{Burgess2003}.  \textit{Right:} Cryo-electron tomograph of one-bound dynein for pre-stroke equilbrium angles \cite{grotjahn2018cryo}.} 
	\label{fig:ParamsPics}
\end{figure}

@@

Despite simulating many combinations of parameters, the number of free variables still produce many degrees of freedom. This makes it difficult to test the entire parameter space and truly invesitage the relationships between different sets of parameters. We approach this problem by individually attributing the parameters to a specific data set from exerpiment. For instance, the exponential unbinding constant, $C$, is fitted specifically to Figure (\ref{fig:trailingbias}). We use time data (\textit{cite paper for time data}\cite{}) to fit the unbinding rate constant, $k_{ub}$, because the inverse yields the average time it takes to unbind, i.e. the both-bound time. The rebinding rates, $k_a$ and $k_b$, are harder to optimize because they directly correlate to step length. The ability for dynein to take long steps is dependent on how long dynein diffuses above the microtubule. Therefore, the rebinding rates have a much larger parameter space due to the flexibility of optimizing according to the one-bound time or step length. Since we are interested in Yildiz's step lengths, we match these rates to the linear regression from Figure (\ref{fig:YildizCorrelation}) and Yildiz's step length plots. The last parameters are the spring constants. The spring constants also influence the step length because they dictate the strength of the restoring force. We fit the spring constants last to control the variance of step lengths.

### 2021-04-27 papers/jin-thesis/Chapters/4Results.tex
THESIS DONE (first fully complete draft)

\begin{table}[H]
  \centering
  \begin{tabular}{|l | r | r | r|}
  	\hline
    Param & Model & Experimental & Source \\
    \hline
    $c_b$ & $\cb \Delta G_{ATP}$ &  & \\
    $c_m$ & $\cm \Delta G_{ATP}$ &  & \\
    $c_t$ & $\ct \Delta G_{ATP}$ &  & \\
    $k_a$ & $\kstk s^{-1}$&  & \\
    $k_b$ & $\kb s^{-1}$&  & \\
    $k_{ub}$ & $\kub s^{-1}$ & & \\
    $C$ & $\cexp$ & & \\
    $L_s$ & $\ls nm$ & $21nm$ & \cite{Burgess2003, 3vkh-cite, carter-paper}\\
    $L_t$ & $\lt nm$ & $23nm$ & \cite{Burgess2003, 3vkh-cite, carter-paper}\\
    $\theta_b$ & $\eqb$ &  120 & \cite{leschziner} \\
    $\theta_m^{\mbox{pre}}$ & $\eqmpre$ &  197 & \cite{Burgess2003}\\
    $\theta_m^{\mbox{post}}$ & $\eqmpost$ & 242 & \cite{Burgess2003}\\
    $\theta_t$ & $\eqt$ &  & \\
    $R_t$ & $\radiust$ & $8nm$ & \cite{Burgess2003}\\
    $R_m$ & $\radiusm$ & $11nm$ & \cite{Burgess2003}\\
    $R_b$ & $\radiusb$ & $3.5nm$ & \cite{Burgess2003}\\
    \hline
  \end{tabular}
  \caption{Optimized model parameters for Figures (\ref{fig:DataFitYildiz}-\ref{fig:Trajectory}).}
  \label{tab:params}
\end{table}

@@

The bottom plot in Figure (\ref{fig:DataFitYildiz}) displays the one-dimensional probability density of step lengths, defined as final displacement ($x_f$) minus initial displacement ($x_i$). The accumulative density from the diagonal shaded region is evident in this plot due to the peak at 0 nm. However, experiment was unable to detect 0 nm steps and is shown within the dip from Experiment Fig 3A (small density at 0 nm were lateral steps on the off-axis with little change on the on-axis). Our model suggests that dynein can still achieve forward directed motion with more positive steps than negatives ones but prefers short stuttered steps with fast one-bound times. A distribution of said one-bound times is shown below in Figure (\ref{fig:OBtime})

@@

As the distribution of times in Figure (\ref{fig:OBtime}) indicates, the model dynein frequently spends $<$ 0.1 $\mu s$ in the one-bound state, with the dynein most likely to rebind immediately after the affinity transitions. This demonstrates the importance of a buffer time when the dynein unbinds in order to avoid instantaneous rebinding and still achieve large positive step lengths. A deeper analysis of this will be discussed in the Limiting Cases section (Section \ref{sec:LimitingCases}).

@@

Based off the resemblence with experimental observations, we are fairly confident with the model's ability to replicate the stepping patterns of real dynein. In our spring model, we assumed each domain to act as springs to influence the powerstroke. However, the simulations displayed more consistent forward stepping when we turned off the binding domain spring and only allowed the motor and tail to have non-zero spring constants. This adjustment displays the flexibility within the model and the potential to fine-tune paramters to fit different sets of data.

\newpage
\section{Limiting Cases}\label{sec:LimitingCases}

As mentioned previously, the model dynein had to take very rapid steps in order to closely generate Yildiz's stepping correlation. To achieve a slope closer to Yildiz's value of 0.6, the binding rate ($k_b$) and affinity transition rate ($k_a$) must both increase significantly. This led to instantaneous rebinding, as we investigated the limiting case of $k_a\gg k_b$ (affinity transition time $t_a\to 0$) and is shown in Figure (\ref{fig:DataFitYildiz99})

\begin{figure}[H]
	\begin{minipage}[b]{0.5\textwidth}
	\includegraphics[width=8.5cm, height=6.5cm]{/mc_plots/u_final_disp_probability_distribution_2.0_5.50e+10_9.00e+99_0.0_1.0_1.0_120.0_197.0_242.0_-0.5}
	\end{minipage}
	\begin{minipage}[b]{0.6\textwidth}		
	\includegraphics[width=9.2cm, height = 6cm]{/mc_plots/u_step_length_1d_probability_density_2.0_5.50e+10_9.00e+99_0.0_1.0_1.0_120.0_197.0_242.0_-0.5}
	\end{minipage}	
	\begin{center}
	\includegraphics[width=12cm, height = 7.5cm]{../../plots/mc_plots/u_ob_time_probability_density_2.0_5.50e+10_9.00e+99_0.0_1.0_1.0_120.0_197.0_242.0_-0.5}
	\end{center}	
	\caption[Limiting Case]{\textbf{Limiting Case for Instant Rebinding.} Each plot closely resembled the plots shown before but with the limiting case of $k_a\gg k_b$. The steps exhibited instantaneous rebinding, where the lifted domain landed at the same position immediately after unbinding off of the microtubule ($x_f\approx x_i$). The one-bound times were recorded within the first time-step where $dt<5$ ns.} 
	\label{fig:DataFitYildiz99}
\end{figure}

When fitting to Yildiz's slope, we avoided instant rebinding by also prioritizing the large positive step lengths Yildiz observed in Experiment Fig 3A (on Figure \ref{fig:DataFitYildiz}). The binding and affinity transition rate can be modified to fit the large step lengths better, but doing so would decrease the correlation between the displacements and thus, the value for the slope. This created a balance between fitting according to the slope or the step length plot. We believe that the parameters shown in Table (\ref{tab:params}) best demonstrates the functionality of the displacements while maintaining dynein's forward directionality with a similar step length trend.

>>>>

\begin{table}[H]
  \centering
  \begin{tabular}{|l | r | r | r|}
  	\hline
    Param & Model & Experimental & Source \\
    \hline
    $c_b$ & $\cb \Delta G_{ATP}$ &  & \\
    $c_m$ & $\cm \Delta G_{ATP}$ &  & \\
    $c_t$ & $\ct \Delta G_{ATP}$ &  & \\
    $k_a$ & $\kstk s^{-1}$&  & \\
    $k_b$ & $\kb s^{-1}$&  & \\
    $k_{ub}$ & $\kub s^{-1}$ & & \\
    $C$ & $\cexp$ & & \\
    $L_s$ & $\ls nm$ & $21nm$ & \cite{Burgess2003, 3vkh-cite, carter-paper}\\
    $L_t$ & $\lt nm$ & $23nm$ & \cite{Burgess2003, 3vkh-cite, carter-paper}\\
    $\theta_b$ & $\eqb$ &  120 & \cite{leschziner} \\
    $\theta_m^{\mbox{pre}}$ & $\eqmpre$ &  197 & \cite{Burgess2003}\\
    $\theta_m^{\mbox{post}}$ & $\eqmpost$ & 242 & \cite{Burgess2003}\\
    $\theta_t$ & $\eqt$ &  & \\
    $R_t$ & $\radiust nm$ & $4nm$ & \cite{Burgess2003}\\
    $R_m$ & $\radiusm nm$ & $5.5nm$ & \cite{Burgess2003}\\
    $R_b$ & $\radiusb nm$ & $1.25nm$ & \cite{Burgess2003}\\
    \hline
  \end{tabular}
  \caption{Optimized model parameters for Figures (\ref{fig:DataFitYildiz}-\ref{fig:Trajectory}).}
  \label{tab:params}
\end{table}

@@

The bottom plot in Figure (\ref{fig:DataFitYildiz}) displays the one-dimensional probability density of step lengths, defined as final displacement ($x_f$) minus initial displacement ($x_i$). The accumulative density from the diagonal shaded region is evident in this plot due to the peak at 0 nm. However, experiment was unable to detect 0 nm steps and is shown from the dip in Experiment Fig 3A (small density at 0 nm were lateral steps on the off-axis with little change on the on-axis). Our model suggests that dynein can still achieve forward directed motion with more positive steps than negatives ones but prefers short stuttered steps with fast one-bound times. A distribution of said one-bound times is shown below in Figure (\ref{fig:OBtime})

@@

As the distribution of times in Figure (\ref{fig:OBtime}) indicates, the model dynein frequently spends $<$ 0.1 $\mu s$ in the one-bound state, with the dynein most likely to rebind immediately after the affinity transitions. This demonstrates the importance of a buffer time when the dynein unbinds to avoid instantaneous rebinding while still achieving large positive step lengths. A deeper analysis of this will be discussed in the Limiting Cases section (Section \ref{sec:LimitingCase}).

@@

Based on the resemblence with experimental observations, we are fairly confident with the model's ability to replicate the stepping patterns of real dynein. In our spring model, we assumed each domain to act as springs to influence the powerstroke. However, the simulations displayed more consistent forward stepping when we turned off the binding domain spring and only allowed the motor and tail to have non-zero spring constants. That is, $c_b=0$ while $c_m=c_t=1$ from Table (\ref{tab:params}). This adjustment displays the flexibility within the model and the potential to fine-tune paramters to fit different sets of data.

\newpage
\section{Limiting Case}\label{sec:LimitingCase}

As mentioned previously, the model dynein had to take very rapid steps in order to closely generate Yildiz's stepping correlation. To lessen the error and achieve a slope closer to Yildiz's value of 0.6, the binding rate ($k_b$) and affinity transition rate ($k_a$) must both increase significantly. This led to instantaneous rebinding, as we investigated the limiting case of $k_a\gg k_b$ (affinity transition time $t_a\to 0$) and is shown in Figure (\ref{fig:DataFitYildiz99})

\begin{figure}[H]
	\begin{minipage}[b]{0.5\textwidth}
	\includegraphics[width=8.5cm, height=6.5cm]{/mc_plots/u_final_disp_probability_distribution_2.0_5.50e+10_9.00e+99_0.0_1.0_1.0_120.0_197.0_242.0_-0.5}
	\end{minipage}
	\begin{minipage}[b]{0.6\textwidth}		
	\includegraphics[width=9.2cm, height = 6cm]{/mc_plots/u_step_length_1d_probability_density_2.0_5.50e+10_9.00e+99_0.0_1.0_1.0_120.0_197.0_242.0_-0.5}
	\end{minipage}	
	\begin{center}
	\includegraphics[width=12cm, height = 7.5cm]{../../plots/mc_plots/u_ob_time_probability_density_2.0_5.50e+10_9.00e+99_0.0_1.0_1.0_120.0_197.0_242.0_-0.5}
	\end{center}	
	\caption[Limiting Case]{\textbf{Limiting Case for Instant Rebinding.} Each plot closely resembled the plots shown before but with the limiting case of $k_a\gg k_b$. The steps exhibited instantaneous rebinding, where the lifted domain landed at the same position immediately after unbinding off of the microtubule ($x_f\approx x_i$). The large majority of one-bound times were recorded within the first time-step where $dt<5$ ns.} 
	\label{fig:DataFitYildiz99}
\end{figure}

When fitting to Yildiz's slope, we avoided instant rebinding by also prioritizing the large positive step lengths Yildiz observed in Experiment Fig 3A (on Figure \ref{fig:DataFitYildiz}). The binding and affinity transition rates can be better optimizied to fit the larger step lengths (we classified ``large'' steps as $> 40$ nm), but doing so would decrease the correlation between the displacements and thus, the value for the slope. This caused an optimization problem of whether to fit according to the slope or the step lengths. We believed that the parameters shown in Table (\ref{tab:params}) best demonstrated the functionality of the displacements from the slope, while maintaining dynein's forward directionality with a similar step length trend towards large steps.

### 2021-04-27 papers/jin-thesis/Chapters/5Conclusion.tex
THESIS DONE (first fully complete draft)

\section{Findings}
\textit{What do our results indicate? Do we think our model correctly represents real dynein and its stepping patterns?}

\section{Further Work}
\textit{How can we improve our model?}

>>>>

%\textit{What do our results indicate? Do we think our model correctly represents real dynein and its stepping patterns?}

We successfully incorporated a Monte Carlo algorithm when dynamically simulating dynein's unique stepping mechanism. By modelling dynein as a two-dimensional particle-rod system, we separated the conformations of the mechanochemical cycle into a both-bound and one-bound state and emulated the powerstroke with springs in each domain. With various degrees of freedom controlling the properties of a step, the model was able to flexibly optimize according to experimental results and provide further conclusions for observations. More specifically, we replicated experimentalist Ahmet Yildiz's unique result of an observered stepping correlation between the initial domain separation and final displacement with a 16.5\% error.

Despite experiments inability to observe step lengths close to 0 nm, we conclude that dynein must take short and fast steps in order to achieve strong stepping dependence. After many tests with different combinations of parameters, the set of parameters shown in Table (\ref{tab:params}) demonstrated Yildiz's stepping correlation while still preserving directed motion with similar positive step length trends. Dynein also preferred larger positive steps when only the motor and tail domains acted as springs. Furthermore, the model achieved dynein's stepping tendencies with biased trailing steps and identical walking trajectories, all while being more computationally efficient and faster than previous simulations.

If time provides, the flexbility of the model allows for a more sufficient optimization process that can minimize the parameter space when replicating experiment. Further work can apply this to other experimental hypotheses regarding dynein's stepping pattern and provide further support for new observations. As mentioned earlier, most simulations model the ATP transitions during the mechanochemical cycle, while this model focuses on the dynamics within the step. Thus, a combined model that incorporates both features may describe natural dynein perfectly and bring us one step closer to engineering a reliable prevention for cell failure.

### 2021-04-27 papers/jin-thesis/main.tex
THESIS DONE (first fully complete draft)

\title{
	\Huge{Modeling the mechanisms of dynein's step using Monte Carlo simulations and Brownian dynamics}\vspace{1em}\\
	\Large{Jin Kiatvongcharoen} \vspace{0.5em}\\[2in] 
%	\includegraphics[width=0.4\columnwidth]{OSU_logo.png}\vspace{1em}\\
	\normalsize{An undergraduate thesis advised by Dr. David Roundy}\\
	\normalsize{submitted to the Department of Physics, Oregon State University}\\%\vspace{2em}\\
	\normalsize{in partial fulfillment of the requirements for the degree of}\\
	\normalsize{B.S. in Physics}\vspace{1.5 em}\\
	\normalsize{Draft submitted \today}\\

@@

	\chapter*{Abstract}
	The motor protein dynein is responsible for cellular processes such as axonal transport and cell division by delivering vital information along the microtubule track. A failed delivery can cause severe damage to the cell’s functionality and lead to neurological diseases. Despite dynein's rich history in cellular research, the mechanisms of its motility remains poorly understood due to its large size, flexible structure, and stochastic stepping. Therefore, recent studies attempts to further understand dynein's stepping mechanism in hopes to prevent motility failure. Experimentalist Ahmet Yildiz and colleagues observed a dependence within dynein's stepping behavior varying as a function of interhead separation. To investigate this, we developed a two-dimensional, particle-rod model of dynein and simulated its stepping mechanisms using Monte Carlo methods and Brownian dynamics. We emulated the powerstroke by simulating rate limiting steps that transition the model dynein between a both-bound and one-bound state. The simulation optimized parameter sets by fitting probability distributions to Yildiz's observations. In order to replicate the correlation observed by Yildiz, the binding rate was increased to encourage fast steps. Our results indicate that dynein takes many short and immediate steps that are undetectable to experimentalists in order to achieve stepping dependence on the initial domain separation. 
%	
%	 However, dynein's walk is unpredictable in nature and can vary in length and direction.
%	
%	 cargo that allows the cell to 
%	Molecular motor proteins are responsible for the cell’s ability to move, divide, and spatially organize itself. A defect within any of the three motor proteins can cause severe damage to the cell’s functionality and lead to neurological diseases. In order to fully understand the cause of such possible diseases, the motor protein dynein and more specifically its structure has been extensively studied. However, there has been limited research focused on the motion of dynein. To investigate this, we will simulate dynein’s stepping mechanisms using Monte Carlo methods and Brownian dynamics. The dynein will be computationally replicated with a particle-rod model undergoing Brownian motion in a fluid, where its stepping will be repeatedly simulated to generate an ensemble of statistics, via the Monte Carlo. The program will optimize dynein’s stepping patterns by fitting parameters according to experimental results of dynein’s step length, time, and probability of stepping.

>>>>

\title{
	\Large{Simulating Dynein: A Monte Carlo Improvement to a Dynamical Model of the Dynein Motor Protein}\vspace{1em}\\
	\large{by} \\ 
	\Large{Jin Kiatvongcharoen} \vspace{0.5em}\\[2in] 
%	\includegraphics[width=0.4\columnwidth]{OSU_logo.png}\vspace{1em}\\
	\normalsize{An undergraduate thesis advised by Dr. David Roundy}\\
	\normalsize{submitted to the Department of Physics, Oregon State University}\\%\vspace{2em}\\
	\normalsize{in partial fulfillment of the requirements for the degree of}\\
	\normalsize{B.S. in Physics}\vspace{1.5 em}\\
	\normalsize{Draft submitted \today}\\

@@

	\chapter*{Abstract}
	The motor protein dynein is responsible for cellular processes such as axonal transport and cell division by delivering vital information along the microtubule track. A failed delivery can cause severe damage to the cell’s functionality and lead to neurological diseases. Despite dynein's rich history in cellular research, the mechanisms of its motility remains poorly understood due to its large size, flexible structure, and stochastic stepping. Unlike its motor protein siblings which posses uniform stepping, dynein's step is unpredictable and can vary in size and direction. However, experimentalist Ahmet Yildiz and colleagues observed a stepping dependence within the interhead separation, indicating a possible correlation during dynein's step. To investigate this, we developed a two-dimensional, particle-rod model of dynein and simulated its stepping mechanisms using Monte Carlo methods and Brownian dynamics. We simplified the mechanochemical cycle by simulating rate limiting steps that transitions between a both-bound and one-bound state and emulated the powerstroke by defining springs in each domain. After optimizing parameters sets, the model achieved Yildiz's stepping correlation only when the binding rates were significantly large. Our simulations indicate that dynein must take many short ($\langle t \rangle < 0.1 \mu s$) and immediate steps, with step lengths close to 0 nm, in order to reproduce a stepping dependence within the interhead domain separation. These results demonstrates the model's ability to capture dynamic properties from experimental observations and provide further support for other experimental hypotheses regarding dynein's step.

%	
%	 However, dynein's walk is unpredictable in nature and can vary in length and direction.
%	
%	 cargo that allows the cell to 
%	Molecular motor proteins are responsible for the cell’s ability to move, divide, and spatially organize itself. A defect within any of the three motor proteins can cause severe damage to the cell’s functionality and lead to neurological diseases. In order to fully understand the cause of such possible diseases, the motor protein dynein and more specifically its structure has been extensively studied. However, there has been limited research focused on the motion of dynein. To investigate this, we will simulate dynein’s stepping mechanisms using Monte Carlo methods and Brownian dynamics. The dynein will be computationally replicated with a particle-rod model undergoing Brownian motion in a fluid, where its stepping will be repeatedly simulated to generate an ensemble of statistics, via the Monte Carlo. The program will optimize dynein’s stepping patterns by fitting parameters according to experimental results of dynein’s step length, time, and probability of stepping.
